---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# MLtoolkit

MLtoolkit is an R package with miscellaneous functions.

## Installation

You can install MLtoolkit from GitHub with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("AndrewKostandy/MLtoolkit")
```

## Example

This is a basic example which shows the truncate_data() function in the package.

Below is a dataframe with numeric columns including univariate outliers:

```{r example}
library(ggplot2)
library(tidyr)
library(MLtoolkit)
theme_set(theme_minimal())

mydata <- tibble(a=(c(10,11,12,seq(70,90,2),50,60)),
                 b=(c(5,11,12,seq(10,20,1),50,60)),
                 c=(c(3,11,12,seq(30,40,1),44,80)),
                 d=(c(0,0,12,seq(20,25,0.5),50,100)))

ggplot(gather(mydata, key, value), aes(key, value)) + geom_boxplot()
```

The truncate_data() function will truncate univariate outliers as follows:

- Values below the 1st quartile by more than 1.5 x IQR are truncated to be exactly 1.5 x IQR below the 1st quartile.

- Values above the 3rd quartile by more than 1.5 x IQR are truncated to be exactly 1.5 x IQR above the 3rd quartile.

After truncation, this is what the data will look like:

```{r example_continued}
mydata2 <- truncate_data(mydata)
ggplot(gather(mydata2, key, value), aes(key, value)) + geom_boxplot()
```

